---
title: "Understanding e-Stat's data model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Understanding e-Stat's data model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
library(estatr)
```

e-Stat trips up first-time users because a "table" is not a flat spreadsheet ŌĆö
it is a cube of coded values plus separate metadata that maps every code to a
label. Understanding that split makes the whole API (and this package) click.

## Tables, values, and classifications

Each table (`statsDataId`) is a set of observations. Every observation is a
single `value` tagged with a code on each of several **classification axes**:

- `tab` ŌĆö the tabulated item (what is being measured, e.g. "population").
- `cat01`..`cat15` ŌĆö category axes (sex, age band, employment status, ...).
- `area` ŌĆö geography (national, prefecture, municipality, or mesh).
- `time` ŌĆö the time period.

The raw data carries only the *codes*. The **metadata** (`getMetaInfo`) maps each
code to a human label. `estat_meta_info()` returns one tibble per axis:

```{r}
meta <- estat_meta_info("0003217721")
names(meta)      # e.g. "tab" "cat01" "cat02" "cat03" "area" "time"
meta$cat01       # code / name / level / parent / unit
attr(meta, "axis_names") # the display name of each axis
```

## How `get_estat()` puts it together

`get_estat()` hides this split. It fetches the data (which conveniently bundles
its own metadata) and performs the codeŌåÆlabel join for you, returning paired
columns:

```{r}
d <- get_estat("0003217721", limit = 100)
# For each axis X you get X (label) and X_code (the raw code)
names(d)
```

Keep the codes when you need stable joins; keep the labels for reading. If you
only want the codes (skipping the metadata join entirely), use the fast path:

```{r}
get_estat("0003217721", decode_labels = FALSE, limit = 100)
```

## Hierarchical codes

Category and area codes are often hierarchical: a `level` column and a `parent`
code let you walk from totals down to sub-categories. For example, an "all ages"
row (`level` 1) may be the parent of individual age bands (`level` 2):

```{r}
meta$cat02[, c("code", "name", "level", "parent")]
```

## Pagination and scale

A single table can be enormous (area-mesh data, long time series). e-Stat caps a
response at 100,000 records; `estatr` pages through the rest automatically,
fetching the remaining pages concurrently by absolute offset. For very large
pulls, pass a `checkpoint` file so an interrupted download resumes where it left
off:

```{r}
big <- get_estat("0003217721", checkpoint = "lfs.checkpoint.rds")
```

## Performance escape hatches

Two arguments skip work you may not need:

- `decode_labels = FALSE` ŌĆö skip the metadata join, return raw codes.
- `as_data_table = TRUE` ŌĆö return the internal `data.table` instead of a tibble,
  avoiding even the boundary conversion for bulk analysis.

Metadata is cached on disk (see `estat_cache_dir()`), so repeated calls for the
same table don't refetch it.
