---
title: "Getting started with estatr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with estatr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  # Chunks that call the e-Stat API need a personal appId and a network
  # connection, so they are not evaluated when the vignette is built.
  eval = FALSE
)
```

`estatr` is a tidy, tidycensus-style interface to the Japanese government-wide
statistics catalog served by the **e-Stat API**. This vignette gets you from a
fresh install to your first labelled table.

## 1. Get an API key

The e-Stat API requires a free application ID (`appId`). Register at
<https://www.e-stat.go.jp/api/> and issue an ID from your mypage (up to three
are allowed per account).

Store it once with `estat_api_key()`:

```{r}
library(estatr)

# Session only
estat_api_key("your-app-id")

# Or persist it to ~/.Renviron for future sessions
estat_api_key("your-app-id", install = TRUE)
```

The key is read from the `ESTAT_API_KEY` environment variable at call time and
is never stored in package state or printed. Treat it as a secret: don't commit
it or paste it into issues. If it leaks, delete and reissue it on your mypage.

## 2. Language

`estatr` works in **English by default** — table names, category labels, area
and time labels, and search all come back in English, because e-Stat itself
provides the translations. The handful of tables with no English release fall
back to Japanese automatically (with a warning). To work in Japanese instead:

```{r}
options(estatr.lang = "J") # for the whole session
# ...or per call: get_estat(id, lang = "J")
```

## 3. Find a table

Every e-Stat table has a `statsDataId`. Search the catalog with
`search_estat()` — English keywords work for the major surveys:

```{r}
tables <- search_estat("Labour Force Survey")
tables[, c("id", "stat_name", "title")]
```

Search coverage is broadest in Japanese (not every table is indexed in
English), so if a search comes up short, try Japanese terms or the curated
shortcuts below.

If you just want a common table without hunting, use a curated shortcut:

```{r}
estat_curated_tables()
lfs <- get_labour_force_survey(limit = 500)
```

## 4. Get labelled data

`get_estat()` is the main entry point. Give it a `statsDataId` and it fetches the
data *and* its metadata in one call, decodes every numeric code to a label, and
returns a tidy tibble:

```{r}
d <- get_estat("0003217721", limit = 500)
d
```

You get one row per observation, with paired label/code columns for each
classification axis (e.g. `area` + `area_code`, `time` + `time_code`), the
`unit`, a numeric `value`, and an `annotation` column that preserves any
non-numeric markers (suppressed cells, footnote symbols) instead of silently
turning them into `NA`. The table's annotation legend is attached as an
attribute:

```{r}
attr(d, "notes")
```

## 5. Reshape and join

Pivot a classification axis into columns with `pivot_estat_wide()`:

```{r}
pivot_estat_wide(d, names_from = "cat01")
```

Join English prefecture names using the bundled `prefectures` table:

```{r}
library(dplyr)
d |>
  left_join(prefectures, by = "area_code")
```

## Where next

- `vignette("finding-tables")` — searching and the curated shortcuts in depth.
- `vignette("time-series")` — working with time series across prefectures.
- `vignette("data-model")` — understanding e-Stat's indicator / classification /
  area / time structure.

## Data source and credit

Statistics are retrieved from the e-Stat API provided by Japan's Statistics
Bureau. Applications that redistribute this data must display the credit line
required by the [e-Stat Terms of Use](https://www.e-stat.go.jp/api/). This
service uses the e-Stat API but its content is not guaranteed by the government.
