---
title: "Finding the right table"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Finding the right table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
library(estatr)
```

The single biggest friction point with e-Stat is finding the `statsDataId` for
the table you actually want. This article covers the three routes `estatr`
gives you: curated shortcuts, friendly search, and the low-level list endpoint.

## Curated shortcuts

For the handful of tables most people want first, there is a named shortcut so
you never touch a `statsDataId`:

```{r}
estat_curated_tables()

get_labour_force_survey(limit = 500)
get_family_income_survey(limit = 500)

# Or by key
get_estat_curated("regional_statistics", limit = 500)
```

Entries whose `statsDataId` is `NA` are recognised names that have not yet been
curated to a specific table вЂ” use search (below) for those, or open an issue to
help curate them.

## Friendly search

`search_estat()` wraps the catalog search and returns a tibble with the most
useful columns renamed to stable snake_case (`id`, `stat_name`, `title`,
`gov_org`, `survey_date`, ...) and moved to the front:

```{r}
# Keyword search (Japanese or English terms)
search_estat("е›Ѕе‹ўиЄїжџ»")

# Combine terms; restrict to recently updated tables
search_estat("еЉґеѓЌеЉ›иЄїжџ» AND йѓЅйЃ“еєњзњЊ", updated_from = "2023")
```

Once you have an `id`, inspect its structure before pulling data (see
`vignette("data-model")`):

```{r}
meta <- estat_meta_info("0003217721")
names(meta) # the classification axes
meta$cat01 # the labels/codes for the first category axis
```

## Low-level listing

For full control over the raw `getStatsList` parameters, use
`estat_stats_list()` directly:

```{r}
estat_stats_list(
  searchWord = "е®¶иЁ€иЄїжџ»",
  statsCode = "00200561",
  limit = 20
)
```

## Discovering downloadable files

Some resources are published as Excel/CSV/PDF rather than as queryable data.
`estat_data_catalog()` finds those:

```{r}
estat_data_catalog(searchWord = "е›Ѕе‹ўиЄїжџ»", limit = 10)
```
