---
title: "Mapping e-Stat data (choropleths)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Mapping e-Stat data (choropleths)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
library(estatr)
```

`estatr` can attach official e-Stat boundary polygons to your data so you can
draw choropleth maps. Geometry comes from e-Stat's own census boundary files,
which are keyed to the **same area codes** as the statistics — so the join is
exact, by construction. This needs the [sf](https://r-spatial.github.io/sf/)
package.

## The one-call path

Pass `geometry = TRUE` to [get_estat()] and you get an `sf` object back:

```{r}
library(sf)
library(ggplot2)

pop <- get_estat(
  "0003433219",             # 2020 Population Census, population by sex
  cdCat01 = "0",            # total (both sexes)
  geometry = TRUE,
  geometry_level = "prefecture",
  geometry_year = 2020
)

ggplot(pop) +
  geom_sf(aes(fill = value)) +
  scale_fill_viridis_c(trans = "log10") +
  labs(title = "Population by prefecture, 2020", fill = "People")
```

## Attaching geometry to data you already have

If you already pulled data, join geometry separately with
[estat_join_geometry()]:

```{r}
d <- get_estat("0003433219", cdCat01 = "0")
sf_d <- estat_join_geometry(d, level = "prefecture", year = 2020)
```

## Levels

`level` (or `geometry_level`) controls the geography:

- `"prefecture"` — 47 prefectures (`area_code` ending in `000`).
- `"municipality"` — cities/wards/towns/villages (5-digit `area_code`).
- `"small_area"` — town-block (町丁・字), the raw 9-digit `KEY_CODE`.
- `"auto"` — prefecture if every `area_code` ends in `000`, else municipality.

Boundaries are downloaded per prefecture and cached (see [estat_cache_dir()]),
so a municipality map of one prefecture only fetches that prefecture's file.

## Matching the boundary year to your data

**This matters for accuracy.** Japanese municipality codes and boundaries change
between censuses (mergers and dissolutions), so set `geometry_year` to the
census year of your data. Joining 2020 data to 2015 boundaries can silently drop
or mis-place municipalities that changed in between.

## Coordinate system

Boundaries come in JGD2000 (`datum = "2000"`, EPSG:4612, the default) or JGD2011
(`datum = "2011"`, EPSG:6668). Reproject as needed for your basemap, e.g. to a
projected CRS for area-correct rendering:

```{r}
pop_proj <- sf::st_transform(pop, 6684) # JGD2011 / Japan Plane Rectangular CS
```

## Designated cities and known limitations

For ordinance-designated cities (政令指定都市) and Tokyo, e-Stat's boundary files
use **ward** codes (e.g. Sapporo's wards `01101`–`01110`), not the parent-city
code (`01100`). Since e-Stat *statistics* also use ward codes for these cities,
the `"municipality"` join is correct row-for-row. But if a particular table
reports a designated city at the parent-city level, that code won't match a
ward polygon — `estat_join_geometry()` warns about any unmatched codes rather
than silently dropping them, so you can spot it. Rolling wards up to parent
cities is a planned convenience.

Note also e-Stat's own disclaimer: these boundaries are drawn for statistical
tabulation and do not necessarily coincide with legal administrative borders.

## Attribution

Boundary data is from e-Stat (統計地理情報システム). Maps you publish must carry
the e-Stat credit line; see the [e-Stat Terms of Use](https://www.e-stat.go.jp/api/).
