---
title: "Working with time series across prefectures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with time series across prefectures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
library(estatr)
library(dplyr)
library(ggplot2)
```

e-Stat's strength is long, consistent time series across all 47 prefectures.
This article shows a typical workflow: pull a series, filter by time and area,
and plot it.

## Filtering server-side

Filter *before* the data leaves e-Stat by passing `cd*` codes to `get_estat()`.
This is far cheaper than pulling everything and filtering in R:

```{r}
# One category, a time window, all areas
d <- get_estat(
  "0003217721",
  cdCat03 = "1",            # e.g. a single sex category
  cdTimeFrom = "2015000000",
  cdTimeTo = "2023000000"
)
```

`cdTime`, `cdArea`, and `cdCat01`..`cdCat15` all accept vectors, which are
comma-joined into e-Stat's code-list form for you:

```{r}
# Just Tokyo and Osaka (whole-prefecture area codes)
get_estat("0003217721", cdArea = c("13000", "27000"))
```

The bundled `prefectures` table has the codes:

```{r}
prefectures |>
  filter(name_en %in% c("Tokyo", "Osaka")) |>
  select(name_en, area_code)
```

## Parsing the time axis

e-Stat time codes are numeric (`time_code`) with a human label (`time`). For
plotting, derive a proper date/period from whichever is convenient. Quarterly
labour-force codes, for example, encode the year in the first four digits:

```{r}
d <- d |>
  mutate(year = as.integer(substr(time_code, 1, 4)))
```

## Plotting

Because `get_estat()` returns a tidy tibble with a numeric `value`, it drops
straight into `ggplot2`:

```{r}
d |>
  filter(!is.na(value)) |>
  ggplot(aes(year, value, colour = cat01)) +
  geom_line() +
  labs(y = unique(d$unit), colour = NULL)
```

## Handling suppressed and annotated values

Non-numeric markers (suppressed cells, footnotes) are *not* silently coerced to
`NA` — they are preserved in the `annotation` column, with the legend attached
as an attribute. Inspect them before dropping:

```{r}
d |> filter(!is.na(annotation)) |> count(annotation)
attr(d, "notes")
```
